Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-24 manuscript/ERL_global_C_review.Rmd
created variables for numbers needed in text, filled in more stats
#29

@@

*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we draw upon *##* records from *##* geographic locations representing *##* C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
These groupings covered *#%* of the primary variable records for forests of known age, or *#%* of total records. 
The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with *#* records total.

We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC  (Table 1). 
For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)). 
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*. 
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand establishment (*i.e.*, that reflected by stand.age). 
Specifically, we removed all plots flagged as managed in ForC-simplified (managed field). 
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.  
Plots flagged as disturbed in ForC-simplified included stands that had undergone anthropogenic thinning or partial harvest (“Cut” or “Harvest” codes) unless this was very minor (percent.mortality= “minor”). 
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods. 
We also removed all plots for which no stand history information had been retrieved.

@@

Of the *#* records in ForC v.#, *#* met the criteria for inclusion in this study (Fig. 1). 
These records were distributed across *#* plots in *#* distinct geographic areas. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests. 
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for *#* fluxes and *#* stocks in tropical broadleaf forests, *#* fluxes and *#* stocks in temperate broadleaf forests, *#* fluxes and *#* stocks in temperate conifer forests, and fluxes and *#* stocks in boreal forests.

@@

**(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
With only # exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both. 
Fluxes tended to be greater in temperate broadleaf than conifer forests, but the difference was never significant.
This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.

@@

There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26). 
Specifically, there were significant differences for only *#* of *#* variables: *LIST*. 
In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest. 
Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC. 
Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.)*

@@

There were only *#* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$. **(more could be added if age trends become significant after outliers are resolved)** 
For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#). 
Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE}
#READING IN TABLES TO GET STATS FOR TEXT

#Table 1 for this manuscript
T1 <- read.csv("tables_figures/C_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
n_records <- sum(T1[,"records"])

#ForC raw
ForC_measurements <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_measurements.csv", stringsAsFactors = FALSE, check.names = FALSE)
ForC_plots <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_plots.csv", stringsAsFactors = FALSE, check.names = FALSE)
ForC_sites <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/data/ForC_sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
n_areas <- max(ForC_sites[,"geographic.area"], na.rm = TRUE )

#ForC simplified
ForC_simplified <- read.csv("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/ForC/ForC_simplified/ForC_simplified.csv", stringsAsFactors = FALSE, check.names = FALSE)
n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
records_eligible_for_inclusion <- NA  #number in ForC_simplified that are not disturbed, managed, or no.history

#stats I haven't yet acquired:
n_records_age_unk <- NA  #total n records no age info
n_records_coord_unk <- NA  #total n records no geographic coordinates

#total numbers of plots and areas (all variables combined)
n_plots_tot <- NA
n_areas_tot <- NA

# percent eligible for inclusion in analysis

n_focal_variable_records_eligible_for_inclusion <- NA #This is prior to assigning biome groupings

# records by biome grouping
n_records_BoB <- NA
n_records_Bomixed <- NA
n_records_Te_mixed <- NA

#numbers of variables with records from ≥7 geographic areas, by biomes/ age class:
n_fluxes_matureTrB_with_ge7_areas <- NA
n_stocks_matureTrB_with_ge7_areas <- NA
n_fluxes_matureTeB_with_ge7_areas <- NA
n_stocks_matureTeB_with_ge7_areas <- NA
n_fluxes_matureTeN_with_ge7_areas <- NA
n_stocks_matureTeN_with_ge7_areas <- NA
n_fluxes_matureBoN_with_ge7_areas <- NA
n_stocks_matureBoN_with_ge7_areas <- NA

n_fluxes_youngTrB_with_ge7_areas <- NA
n_stocks_youngTrB_with_ge7_areas <- NA
n_fluxes_youngTeB_with_ge7_areas <- NA
n_stocks_youngTeB_with_ge7_areas <- NA
n_fluxes_youngTeN_with_ge7_areas <- NA
n_stocks_youngTeN_with_ge7_areas <- NA
n_fluxes_youngBoN_with_ge7_areas <- NA
n_stocks_youngBoN_with_ge7_areas <- NA

```

@@

*Methods/Design.* Here, we draw from the Global Forest C Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we draw upon `r n_records` records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

@@

Analyses drew from ForC-simplified (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which is a rearrangement of ForC intended to facilitate analyses. 
In generating ForC-simplified, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ (IPCC 2006). 
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database). 
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event. 
Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%). 
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone anthropogenic thinning or partial harvest unless this was very minor. 
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/nrow(ForC_simplified)*100,digits=1)`%). 
In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.

We selected 23 annual flux and 11 C stock variables for inclusion in the analysis, although two flux variables ($R_{het-ag}$ and $R_{het}$) were included for conceptual completeness but had no records in ForC (Table 1). 
Records for these variables represented `r round(n_focal_variable_records_eligible_for_inclusion/records_eligible_for_inclusion*100,digits=1)`% of the total records eligible for inclusion.
For this analysis, we combined some of ForC’s specific variables (*e.g.*, multiple variables for net primary productivity including various components) into more broadly defined variables (Tables 1, S#(variable mapping)). 
Throughout ForC, for all measurements drawing from tree census data (*e.g.*, biomass, productivity), the minimum stem diameter sampled was $\le$ 10cm. All records were measured directly or derived from field measurements (as opposed to modeled).

@@

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Assigning stands to these groupings required the exclusion of records for which ForC lacked records of stand age (n=`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of total database), or geographic coordinates (n=`r round(n_records_coord_unk/nrow(ForC_measurements)*100, digits=1)`% of total database). 
In total, our analysis retained `r round(n_records/n_focal_variable_records_eligible_for_inclusion*100, digits=1)` of the focal variable records for forests of known age. 
The most well-represented forest types excluded were boreal broadleaf (`r n_records_BoB` records) and boreal and temperate mixed broadleaf-needleleaf (`r n_records_Bomixed` and `r n_records_Te_mixed` records, respectively).

@@

Of the `r nrow(ForC_measurements)` in ForC v.#, `r n_records` met the criteria for inclusion in this study (Fig. 1). 
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas. 
Of the 23 flux and 11 stock variables mapped in these diagrams, ForC contained mature forest estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_matureTrB_with_ge7_areas` fluxes and `r n_stocks_matureTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_matureTeB_with_ge7_areas` fluxes and `r n_stocks_matureTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_matureTeN_with_ge7_areas` fluxes and `r n_stocks_matureTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_matureBoN_with_ge7_areas` fluxes and `r n_stocks_matureBoN_with_ge7_areas` stocks in boreal forests. 
For regrowth forests (<100 yrs), ForC contained estimates from $\ge$ 7 distinct geographic areas for `r n_fluxes_youngTrB_with_ge7_areas` fluxes and `r n_stocks_youngTrB_with_ge7_areas` stocks in tropical broadleaf forests, `r n_fluxes_youngTeB_with_ge7_areas` fluxes and `r n_stocks_youngTeB_with_ge7_areas` stocks in temperate broadleaf forests, `r n_fluxes_youngTeN_with_ge7_areas` fluxes and `r n_stocks_youngTeN_with_ge7_areas` stocks in temperate conifer forests, and `r n_fluxes_youngBoN_with_ge7_areas` fluxes and `r n_stocks_youngBoN_with_ge7_areas` stocks in boreal forests.

@@

**(The following covers only the variables with age trends figures; there maybe a few more without much age data)**
With only ## exceptions, C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests (Table 1, Figs. 6, S1-S15).
In this majority of cases, differences between tropical and boreal forests were always significant, with temperate forests intermediate and significantly different from one or both. 
Fluxes tended to be greater in temperate broadleaf than conifer forests, but the difference was never significant.
This pattern held for the following variables: $GPP$, $NPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{foliage}$, $ANPP_{litterfall}$, $BNPP$, $BNPP_fine$, $R_{eco}$, $R_{root}$, $R_{soil}$, and $R_{het-soil}$.

@@

There were fewer distinct trends in C stocks across biomes (Figs. 7, S16-S26). 
Specifically, there were significant differences for only *##* of *##* variables: *LIST*. 
In all of these cases, temperate conifer forests were in the highest significance grouping, and boreal forests in the lowest. 
Tropical and temperate broadleaf forests fell in between--most commonly being significantly different from temperate conifer but not boreal forests.
Once again, the high values for the temperate conifer biome were driven by the very high-biomass forests of the US Pacific Northwest, which are disproportionately represented in the current version of ForC. 
Thus, biome differences should be interpreted more as driven more by geographic distribution of sampling than by true differences.)*

@@

There were only *##* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$. **(more could be added if age trends become significant after outliers are resolved)** 
For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#). 
Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).
